---
title: "Assignment 1"
author: "PD, JO, and MvD, group 06"
date: "17 febuary 2023"
output: pdf_document
highlight: tango
---

## Exercise 1

Here is the part for exercise 1

**a)**

```{r}
# what is code without comments
print("exercise 1.a code here!")

```

## Exercise 2

```{r}
fat = scan("../datasets/cholesterol.txt", what = list(before = 0, after = 0)) 
attach(fat)
```

**a)**

To investigate the data set we create a box plot of both columns. Judging from this we can observe a possible difference. the mean of the data from after 8 weeks appears to be lower.

```{r}
boxplot(fat)
```

Next we plot a normal Q-Q plot to check if the data is normally distributed. this appears to be the case.

```{r}
par(mfrow = c(1,2)); qqnorm(before); qqnorm(after)
```

To check if the before and after 8 weeks data is correlated we can plot the two data sets against each other. The plot shows a clear linear correlation between before and after. Then we can confirm this with both the Pearson's and Spearman's correlation test. Both of these give the conclusion that there indeed is a correlation and it is a strong correlation with **an r value of 0.99.**

```{r}
plot(before~after)

cor.test(before, after)
cor.test(before, after, method="spearman")
```

**b)**  Apply two relevant tests (cf. Lectures 2, 3) to verify whether
the diet with low fat margarine has an effect (argue whether the data
are paired or not). Is a permutation test applicable?

Since the data is collected from the same patient before and after 8 weeks it is clear the data is paired. We can therefore perform a t-test to investigate whether there is a difference in the mean of the before and after data.

```{r}
t.test(before, after, alternative="greater", paired = TRUE)
```

This shows that there is an actual difference in mean between the before and after data.

We also perform a Wilcoxon signed rank test to check if the same conclusion will be drawn. This test concludes the same. There is a significant difference in mean between the before and after data.

```{r}
wilcox.test(before, after, alternative = "greater")
```

```{r}
permStat = function(x, y) { mean(x-y) }

B = 1000; tstar=numeric(B);
i =0
for(i in i:B) {
  cbind = cbind(before, after)
  fatStar = t(apply(cbind,1,sample))
  tstar[i] = permStat(fatStar[,1], fatStar[,2]) 
}

t = permStat(before, after)
print(t)

hist(tstar, xlim = c(-0.6,0.6))
lines(rep(t,2), c(0,20), col = "red", lwd=2)
p1=sum(tstar<t)/B
pr=sum(tstar>t)/B
p=2*min(p1,pr); p
#conclusion indeed significant difference

```

**c)**

```{r}
# bootstrap test to find out if fat$after is uniformly distributed. and construct a 95% confidence interval. (Kolmogorov-Smirnov?)

```

**d)**

```{r}

```

## Exercise 3

Here is the part for exercise 3 **a)**

```{r}
# what is code without comments
print("exercise 3.a code here!")

```

## Exercise 4

Here is the part for exercise 4 **a)**

```{r}
# what is code without comments
print("exercise 4.a code here!")

```
