---
title: "Assignment 1"
author: "PD, JO, and MD, group 70"
date: "17 febuary 2023"
output: pdf_document
highlight: tango
---

## Exercise 1

Here is the part for exercise 1

```{r}
birthweight = read.table(file="../datasets/birthweight.txt", header=FALSE)
birthweight <- birthweight[2:189,]
birthweight <- as.numeric(unlist(birthweight))
```

**a)**

```{r}
# check normality
shapiro.test(birthweight)
qqnorm(birthweight)
 
# 96%-CI
t.test(birthweight, mu=0, conf.level = 0.96)

#sample size needed for CI length of 100
e = 100/mean(birthweight)
qnorm(0.98)^2*0.96*0.04/(e/2)^2

#paul had de bootstrap
```

**b)**

```{r}
#t-test to verify mean weight is bigger than 2800
t.test(birthweight, mu=2800, alternative = "g")

#sign test
binom.test(sum(birthweight>2800), length(birthweight), alternative = "g")
```

**c)**

```{r}
#geen idee of dit goed is
#power of t-test and sign test
B=10000; n=188; s = sd(birthweight); m = mean(birthweight)
psign = numeric(B)
pttest = numeric(B)

for(i in 1:B){
  x = rnorm(n, mean=m, sd=s)
  pttest[i] = t.test(x, mu=2800, alternative = "g")[[3]]
  psign[i] = binom.test(sum(x>2800), n, alternative ="g")[[3]]
}

sum(pttest<0.05)/B
sum(psign<0.05)/B
```

**d)**

```{r}
n = length(birthweight)
lower_bound = 0.25
p_hat = sum(birthweight<2600)/n
margin = p_hat - lower_bound
upper_bound = lower_bound + 2*margin
z = margin / sqrt((p_hat*(1-p_hat))/n)
alpha = (1 - pnorm(z))*2
confidence_level = 1 - alpha
```

**e)**

```{r}

```

## Exercise 2

```{r}
fat = scan("../datasets/cholesterol.txt", what = list(before = 0, after = 0)) 
attach(fat)
```

**a)**

To investigate the data set we create a box plot of both columns. Judging from this we can observe a possible difference. the mean of the data from after 8 weeks appears to be lower.

```{r}
boxplot(fat)
```

Next we plot a normal Q-Q plot to check if the data is normally distributed. this appears to be the case.

```{r}
par(mfrow = c(1,2)); qqnorm(before); qqnorm(after)
```

To check if the before and after 8 weeks data is correlated we can plot the two data sets against each other. The plot shows a clear linear correlation between before and after. Then we can confirm this with both the Pearson's and Spearman's correlation test. Both of these give the conclusion that there indeed is a correlation and it is a strong correlation with **an r value of 0.99.**

```{r}
plot(before~after)

cor.test(before, after)
cor.test(before, after, method="spearman")
```

**b)**  Apply two relevant tests (cf. Lectures 2, 3) to verify whether the diet with low fat margarine has an effect (argue whether the data are paired or not). Is a permutation test applicable?

Since the data is collected from the same patient before and after 8 weeks it is clear the data is paired. We can therefore perform a t-test to investigate whether there is a difference in the mean of the before and after data.

```{r}
t.test(before, after, alternative="greater", paired = TRUE)
```

This shows that there is an actual difference in mean between the before and after data.

We also perform a Wilcoxon signed rank test to check if the same conclusion will be drawn. This test concludes the same. There is a significant difference in mean between the before and after data.

```{r}
wilcox.test(before, after, alternative = "greater")
```

```{r}
permStat = function(x, y) { mean(x-y) }

B = 1000; tstar=numeric(B);
i =0
for(i in i:B) {
  cbind = cbind(before, after)
  fatStar = t(apply(cbind,1,sample))
  tstar[i] = permStat(fatStar[,1], fatStar[,2]) 
}

t = permStat(before, after)
print(t)

hist(tstar, xlim = c(-0.6,0.6))
lines(rep(t,2), c(0,20), col = "red", lwd=2)
p1=sum(tstar<t)/B
pr=sum(tstar>t)/B
p=2*min(p1,pr); p
#conclusion indeed significant difference

```

**c)**

```{r}
# bootstrap test to find out if fat$after is uniformly distributed. and construct a 95% confidence interval. (Kolmogorov-Smirnov?)

```

**d)**

```{r}

```

## Exercise 3

```{r}
diet = read.table("../datasets/diet.txt", header = TRUE)
```

Here is the part for exercise 3 **a)**

```{r}
# what is code without comments

```

**b)**

```{r}

```

## Exercise 4

Here is the part for exercise 4 **a)**

```{r}
# what is code without comments
print("exercise 4.a code here!")

```
