---
title: "Assignment 1"
author: "PD, JO, and MvD, group 06"
date: "17 febuary 2023"
output: pdf_document
highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

## Exercise 1

Here is the part for exercise 1 **a)**

```{r}
# what is code without comments
print("exercise 1.a code here!")

```

**b)** **c)**

**e)**

-   Null hypothesis (H0): The mean weight is the same for male and
    female babies.

-   Alternative hypothesis (Ha): The mean weight is different for male
    and female babies.

Our p-value of 0.56 is higher then alpha of 0.05 this means we cant
reject our null hypothesis

```{r}
male_below_2600 = 34
female_below_2600 = 28

male_total = 95
female_total = 91

# Create vectors of the data
prop.test(c(male_below_2600,female_below_2600), c(male_total,female_total))
```

## Exercise 2

Here is the part for exercise 2 **a)**

```{r}
# what is code without comments
print("exercise 2.a code here!")

```

## Exercise 3

Here is the part for exercise 3 **a)**

```{r}
# what is code without comments
diet <- read.table("../datasets/diet.txt", header = TRUE)


par(mfrow=c(1,2))
boxplot(x = diet$weight6weeks,xlab= "Weight 6 weeks" )
boxplot(x = diet$preweight,  xlab= "Pre-weight")

#Check for if data is normaly distributed
hist(diet$weight6weeks - diet$preweight, main = "Difference in weight after 6 weeks", xlab = "Difference")
qqnorm(diet$weight6weeks - diet$preweight)
qqline(diet$weight6weeks - diet$preweight)


t.test(diet$preweight, diet$weight6weeks, paired = TRUE)
```

## Exercise 4

Here is the part for exercise 4 **a)**

```{r}
# what is code without comments
print("exercise 4.a code here!")

```
