---
title: "Assignment 1"
author: "PD, JO, and MD, group 70"
date: "17 febuary 2023"
output: pdf_document
highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

## Exercise 1

**a)**

```{r}
birthweight = read.table(file="../datasets/birthweight.txt", header=FALSE)
birthweight <- birthweight[2:189,]
birthweight <- as.numeric(unlist(birthweight))
```

-   Null hypothesis (H0): The mean weight is the same for male and
    female babies.

-   Alternative hypothesis (Ha): The mean weight is different for male
    and female babies.

Our p-value of 0.56 is higher then alpha of 0.05 this means we cant
reject our null hypothesis

```{r}
# check normality
shapiro.test(birthweight)
qqnorm(birthweight)
 
# 96%-CI
t.test(birthweight, mu=0, conf.level = 0.96)

#sample size needed for CI length of 100
e = 100/mean(birthweight)
qnorm(0.98)^2*0.96*0.04/(e/2)^2

#paul had de bootstrap
```

**b)**

```{r}
#t-test to verify mean weight is bigger than 2800
t.test(birthweight, mu=2800, alternative = "g")

#sign test
binom.test(sum(birthweight>2800), length(birthweight), alternative = "g")
```

**c)**

```{r}
#geen idee of dit goed is
#power of t-test and sign test
B=10000; n=188; s = sd(birthweight); m = mean(birthweight)
psign = numeric(B)
pttest = numeric(B)

for(i in 1:B){
  x = rnorm(n, mean=m, sd=s)
  pttest[i] = t.test(x, mu=2800, alternative = "g")[[3]]
  psign[i] = binom.test(sum(x>2800), n, alternative ="g")[[3]]
}

sum(pttest<0.05)/B
sum(psign<0.05)/B
```

**d)**

```{r}
n = length(birthweight)
lower_bound = 0.25
p_hat = sum(birthweight<2600)/n
margin = p_hat - lower_bound
upper_bound = lower_bound + 2*margin
z = margin / sqrt((p_hat*(1-p_hat))/n)
alpha = (1 - pnorm(z))*2
confidence_level = 1 - alpha
```

**e)**

To test the claim that there is a difference in the mean birth weight of
male and female babies are different we can perform a prop test on the
proportions of males and females that are born with a weight below 2600
gram.

-   H_null : There is no difference in proportion between male and
    female babies born with a weight less than 2600.

-   H_alt : There is a difference in proportion between male and female
    babies born with a weight less than 2600.

This test concludes with a p-value of \~0.5 and therefore we cannot
reject the null hypothesis.

We can not directly reject the claim that there is a difference in the
mean birth weight between male and female babies because we cannot
directly measure this. However, we also have not found evidence for it
in the given proportions.

```{r}
male_weights <- c(34, 61)
female_weights <- c(28, 65)

x<-c(male_weights[1],female_weights[1])
y<-c(sum(male_weights),sum(female_weights))

prop.test(x, y)
```

## Exercise 2

```{r}
fat = scan("../datasets/cholesterol.txt", what = list(before = 0, after = 0));
attach(fat);
```

**a)**

To investigate the data set we create a box plot of both columns.
Judging from this we can observe a possible difference. the mean of the
data from after 8 weeks appears to be lower.

```{r}
boxplot(fat)
```

Next we plot a normal Q-Q plot to check if the data is normally
distributed. this appears to be the case.

```{r}
par(mfrow = c(1,2)); qqnorm(before); qqnorm(after)
```

To check if the before and after 8 weeks data is correlated we can plot
the two data sets against each other. The plot shows a clear linear
correlation between before and after. Then we can confirm this with both
the Pearson's and Spearman's correlation test. Both of these give the
conclusion that there indeed is a correlation and it is a strong
correlation with **an r value of 0.99.**

```{r}
plot(before~after)

cor.test(before, after)
cor.test(before, after, method="spearman")
```

**b)**

Since the data is collected from the same patient before and after 8
weeks it is clear the data is paired. We can therefore perform a t-test
to investigate whether there is a difference in the mean of the before
and after data.

-   H_null : mean(before) = mean(after)
-   H_alt : mean(before) \> mean(after)

Performing a t.test gives us a p-value of 1.639e-11 this is smaller than
our significance level of 0.05 and we can therefore reject the null
hypothesis. The before data has a mean that is greater than the mean of
the after data.

```{r}
t.test(before, after, alternative="greater", paired = TRUE)
```

We also perform a Wilcoxon signed rank test to check if the same
conclusion can be drawn. The test gives us a p-value of 3.815e-06. with
significance level of 0.05 we can reject the null hypothesis again.

```{r}
wilcox.test(before, after, alternative = "greater", paired = TRUE)
```

A permutation test is applicable for independent samples with any test
statistic that expresses difference between the samples. This means that
for the before and after data we can perform this test is the data is
independent. The before and after data are independent therefore we
might perform a permutation test.

**c)**

The estimator of theta we find by computing max(after). Now we can find
the confidence interval for this estimator by the bootstrapped
confidence interval method. This gives us a confidence interval of
[7.67, 8.38]

```{r}
B=1000
T1 = max(after)
Tstar=numeric(B)
c1 = after

for(i in 1:B) {
  Xstar=sample(c1,replace=TRUE)
  Tstar[i]=max(Xstar)
}
Tstar25=quantile(Tstar,0.025)
Tstar975=quantile(Tstar,0.975)

c(2*T1-Tstar975,2*T1-Tstar25)
```

**d)**

To determine for which thetas we cannot reject that our estimate

-   h0: p(after) = unif(3,theta) where theta in [3, 12]

-   h1: p(after) =\\= unif(3,theta) where theta in [3, 12]

We performed the bootstrap test for the values for theta in 3 ... 12.
For the values inside the bootstrapped confidence interval we cannot
reject the null hypothesis.

the Kolmogorov-Smirnoff (KS) test can be used to perform this analysis.
This can be done by generating data from the uniform distribution and
performing the KS test on the generated data and the after data.

```{r}
data = after
n=length(data); t=max(data); t

B=1000; tstar=numeric(B)
for (i in 1:B) {
  xstar=runif(n, 3, 9)
  tstar[i]=max(xstar)
}

pl=sum(tstar<t)/B;
pr=sum(tstar>t)/B
p=2*min(pl,pr); p

```

**e)**

Using the sign test we are test if H0: median(after) = 6. with H1:
median(after ) \< 6. Performing the test gives us a p-value of 0.61111
this is not lower than our significance level of 0.05 and therefore we
cannot reject the null hypothesis.

```{r}
x = sum(after<6)
n = length(after<6)
binom.test(x, n, x/n, "l")
```

Next to the wilcox.

```{r}
?wilcox.test(after,mu=6)
```

## Exercise 3

```{r}
diet = read.table("../datasets/diet.txt", header = TRUE)
```

Here is the part for exercise 3 **a)**

```{r}
par(mfrow=c(1,2))
boxplot(x = diet$weight6weeks,xlab= "Weight 6 weeks" )
boxplot(x = diet$preweight,  xlab= "Pre-weight")

#Check for if data is normaly distributed
hist(diet$weight6weeks - diet$preweight, main = "Difference in weight after 6 weeks", xlab = "Difference")
qqnorm(diet$weight6weeks - diet$preweight)
qqline(diet$weight6weeks - diet$preweight)


t.test(diet$preweight, diet$weight6weeks, paired = TRUE)
```

**b)**

Our goal is to use one-way ANOVA test to compare the diets. We first
draw QQ-plot to check if we can assume normality this seems to be the
case so go over to our hypothesis.

-   Null hypothesis: The means of the lost weight are equal among the
    three types of diets.

-   Alternative hypothesis: The means of the lost weight are not equal
    among the three types of diets.

    Our p-value of 0.832 is greater than our significance level of 0.05
    this means we can not reject our null hypothesis\
    The Kruskal-Wallis is also valid because it has the same
    pre-conditions as the anova test the difference is that
    Kruskal-Wallis test can also be used when normality fails

```{r}
dietframe <- data.frame(weight=diet$weight6weeks, diet=factor(diet$diet))
dietanov=lm(weight~diet ,data = dietframe)
anova(dietanov)
par(mfrow=c(1,2)); qqnorm(residuals(dietanov))
qqline(residuals(dietanov))
plot(fitted(dietanov),residuals(dietanov))
```

**c )**

Our goal is to use two-way ANOVA test to compare the diets. We first
draw QQ-plot to check if we can assume normality this seems to be the
case so we can go over to our hypothesis.

-   Null hypothesis: The means of the lost weight are equal among the
    three types of diets.

-   Alternative hypothesis: The means of the lost weight are not equal
    among the three types of diets.

Our p-value based on interaction between the parameters is 0.758 is
greater than our significance level of 0.05 this means we can not reject
our null hypothesis. There seems to be no interaction between the
interaction those variables have on response variable. Although gender
seems to have an effect on weight loss.

```{r}
dietframe <- data.frame(weight=diet$weight6weeks, diet=factor(diet$diet), gender=factor(diet$gender))
dietanov=lm(weight~diet*gender ,data = dietframe)
anova(dietanov)
par(mfrow=c(1,2))
qqnorm(residuals(dietanov))
plot(fitted(dietanov), residuals(dietanov))
```

**d ) Dropped**

**e )**

⚠️**Ome Paul kijk hier nog even na ⚠️**

We prefer the use of the one-way anova-test, because it focusses on
relation between one two variables which makes more clear to spot a
relation.

## Exercise 4

Here is the part for exercise 4 **a)**

```{r}
# what is code without comments
print("exercise 4.a code here!")
# 
```
